Comparison of MCMC algorithms in Stochastic Volatility Models

### 2023-08-30 writeups/proposal_presentation/proposal_presentation_sv.qmd
reformat

## Background
- Variety of Bayesian estimation strategies for different types of Stochastic Volatility models using classical Markov Chain Monte Carlo (MCMC) algorithms

@@

- However, as the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms

- How do we ensure the validity of inferences from our algorithms across different estimation strategies?

@@

:::{.nonincremental}
- $y_t$ is mean corrected returns for equally spaced intervals $t$

@@

@@

## Simulation Based Calibration (SBC)
<!-- Probably need to talk about what calibration means-->
- The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.

@@

- SBC compares the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.

- If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage. That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval.

@@

## Simulation Based Calibration (SBC)
A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$. This is implied by the following expression (and proven in Talts et al. (2018)):

$$
\begin{aligned}
\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
\end{aligned}
$$

That is, the posterior averaged over the joint distribution follows the same distribution as the prior. Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.

@@

## HMC Results

![1000 Iterations of SBC (HMC)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=80%}

## HMC Results

![5000 Iterations of SBC (HMC)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){width=75%}

## Gaussian mixture Results

![1000 Iterations of SBC (KSC)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}

## Next steps

- Increasing number of posterior draws for KSC model
    + MCMC may not have converged onto target distribution

>>>>

## Background
- Stochastic volatility models the variance of a financial instrument as a random variable

- Variety of Bayesian estimation strategies using classic Markov Chain Monte Carlo (MCMC) algorithms such as Gibbs and Metropolis Hastings

@@

- As the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms. How do we ensure the validity of our inferences across different estimation strategies?

@@

<!-- :::{.nonincremental} -->
- $y_t$ is mean corrected returns for equally spaced intervals $t$

@@

<!-- ::: -->

@@

## Simulation Based Calibration (SBC)
<!-- Probably need to talk about what calibration means-->
<!-- - The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.

@@

If B computation is correct it can capture the known parameers.-->
:::{.nonincremental}
SBC simulates data with known parameters and checks if on average, the interval of an estimated model contains the true parameter

1) Take a draw from the joint prior $\theta^{sim}$
2) Generate a dataset conditional on the prior draw
3) Calculate a rank statistics
4)  Compare the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.

If Bayesian computation is well calibrated, then the posterior (credible) intervals will have sufficient coverage. That is, for any percentage interval selected (for example 90%) then 90\% of these constructed intervals will contain $\theta^{sim}$.

@@

## Simulation Based Calibration (SBC)
A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples.

$$
\begin{aligned}
\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} 
\end{aligned}
$$
<!-- &= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
 -->

That is, the posterior averaged over the joint distribution follows the same distribution as the prior. The rank statistic for a given parameter and simulation follows a discrete uniform distribution (proven in Talts et al. (2018)).

$$
\begin{aligned}
r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
\end{aligned}
$$

Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.

@@

## HMC Results (1000 iterations)

![](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){fig-align="center" height=90%}

## HMC Results (5000 iterations)

![](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){fig-align="center" height=90%}

## Gaussian mixture Results (1000 iterations)

![](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){fig-align="center" height=90%}

## Next steps
:::{.nonincremental}
- Increasing number of posterior draws for KSC model
    + MCMC may not have converged onto target distribution
